---
title: "Aides Drug"
author: "Monserrat Mendez Mora"
date: "Spring 2024"
output: html_document
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
triUmp <-read.csv(file="triUmpsal.csv",header=TRUE)
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

A new  preventative aides drug (triUmpsal) is under test in clinical trials. Consider only the DrugUsers, does this drug seem effective?  Hint - use dplyr to obtain that subpopulation with the filter command.


```{r}
drugies <- filter(triUmp, DrugUser == "yes")
```




# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->


```{r}
barchartGC(~infected + treatment,data=drugies)
barchartGC(~infected + treatment,data=drugies, type="percent")

```

The data presented in the graph leaves no doubt that the drug is highly effective in preventing AIDS. As the graph clearly shows, that almost 80% of patients who reported having AIDS had taken a placebo instead of the actual drug. This striking evidence reinforces the conclusion that the drug is a crucial tool in preventing the spread of AIDS.

### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->


```{r}
table1 <- xtabs(~infected + treatment,data=drugies)
rowPerc(table1)
colPerc(table1)
```

The number of people who reported getting aids in the placebo group was three times higher than those who were taking the actual drug.



## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->


```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```
  In this section, we conducted three inferential tests, consisting of two Chi-squared tests and a Fisher's Exact Test. The first Chi-squared test resulted in a p-value of 0.0000001242, which was well below the significance level of 0.05. The second Chi-squared test produced a value of 27.955 (approximately 28), which exceeded the critical value for the degrees of freedom in this case, which was 1. Based on these results, we can conclude that there is a significant difference in the data.  


# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->


From our graphical and numerical results, it is evident that the frequency of disease contraction was higher in the placebo group as compared to the treatment group. Moreover, the statistical inference test shows significant differences in the data collected from both groups. Based on these findings, we can conclude that there is a negative correlation between the use of the drug and the incidence of disease. Since a large number of participants were included in the study, we can assume that random variation has a negligible impact on the data. Therefore, we can infer from the results, assuming the data is accurate, that the drug use has a positive effect, and it can be considered as an effective treatment for the disease.


